---
title: "30-feature-engineering"
output: html_notebook
---


## Just as a reminder ~ Predictive Modeling Problem:

We want to predict if a person is going to score higher in **neuroticism** based on how they score in other dimensions with respect to their **age group** or **gender identity**. 

**Score Calculation**
The first thing to do here would be calculate the scores for the big five dimensions.

Calculation the Extraversion, Aggreableness, Neuroticism, Conscientiousness and Openness to experience, scores. 

The score determine which range you are in.

The test has 10 questions for each trait. (For Extraversion - E1:E10, Agreeableness - A1:A10, etc) Each question has a five point scale going from "0 -  strongly disagree" to "5 -  strongly agree".

These questions would either favor the specfic personality trait, or go against it. In my analysis, I chose to combine the scores from each dimension into a single dimension and then added or substracted the score from a single question, based on if the question supports or goes against a particular trait. 

for example:

E1	I am the life of the party. - Score from this question would be added
E2	I don't talk a lot. - Score from this question would be substracted

Let's create a vector with all the column names for the dimensions
```{r dimensions vector}

dimensions <- c("extraversion", "neuroticism", "agreeableness", "conscientiousness", "openness")

```


#Creating dimensional scores
```{r Creation of Dimensional Score}
get_scores <- function(df){

  df %>% 
    mutate(extraversion = e1 - e2 + e3 - e4 + e5 - e6 + e7 - e8 + e9 - e10, 
           neuroticism = n10 + n9 + n8 + n7 + n6 + n5 - n4 + n3 - n2 + n1, 
           agreeableness = a10 + a9 + a8 - a7 + a6 - a5 + a4 - a3 + a2 - a1, 
           conscientiousness = c10 + c9 - c8 + c7 - c6 + c5 - c4 + c3 - c2 + c1, 
           openness = o10 + o9 - o8 + o7 - o6 + o5 - o4 + o3 - o2 + o1) %>% 
    select(extraversion, neuroticism, agreeableness, conscientiousness, openness,  everything())
}

```

```{r Checking Dimensional Scores for NAs}
#Verify that there are no NA's
personas %>% 
  get_scores() %>% 
  verify(not_na(dimensions))
```
# [MISSING FILL THIS IN]
#Normalization
We will need to normalize these as all of them have different ranges. 

```{r Normalization By Variable Function}
normalization <- function(x, na.rm = T) (x - min(x, na.rm = T))/(max(x, na.rm = T)-min(x, na.rm = T))
#[Code here for normalized function]
norm_by_var <- function(df, vars)
{
  df %>% 
    mutate_at(vars, list(norm = normalization)) %>% 
    select(contains('_norm'), vars, everything())
}

#[Code here that verifies normalized values created are within the bounds of 0 and 1]
get_scores(personas) %>%
  norm_by_var(dimensions) %>%
  assert(within_bounds(0,1), extraversion_norm) %>%
  assert(within_bounds(0,1), neuroticism_norm) %>%
  assert(within_bounds(0,1), agreeableness_norm) %>%
  assert(within_bounds(0,1), conscientiousness_norm) %>%
  assert(within_bounds(0,1), openness_norm)
```

#Creating age brackets

We will need to split up the age in a way that is reasonable. Let's first look at the distribution of the age of the people who took the personality assessment. 

```{r Creating an age bracket}
personas%>% 
  ggplot(aes(x = age))+
  geom_histogram()

personas%>% 
  group_by(age) %>% 
  summarize(n=n()) %>% 
  ungroup() %>% 
  arrange(age, -n)
```
From here we can see that the number of observations or subjects of age 53 and older is going to be lower than 100. I think this will reduce the quality of summary measures that we can derive from the splits.
For this reason, **I will cut off the ages past 60.**

```{r}
personas <- personas%>% 
  filter(age<=60)
```


The brackets I will create will be:
Bracket 1: 13-18 - Teenasge years
Bracket 2: 19-21 - Early adulthood
Bracket 3: 22-25 - Early 20s
Bracket 4: 26-29 - later 20s
Bracket 5: 30-34 - Early 30s
Bracket 6: 35-39 - Late 30s
Bracket 7: 40-44 - Early 40s
Bracket 8: 45-49 - Late 40s
Bracket 9: 50-60 - 50s 

This provides age cuts which are meaningful enough. 

```{r Age Bracket Function}

bracket_creator <- function(x){
  if (x >= 13 & x <= 18){"Bracket1"}
  else if (x >= 19 & x <= 21){"Bracket2"}
  else if (x >= 22 & x <= 25){"Bracket3"}
  else if (x >= 26 & x <= 29){"Bracket4"}
  else if (x >= 30 & x <= 34){"Bracket5"}
  else if (x >= 35 & x <= 39){"Bracket6"}
  else if (x >= 40 & x <= 44){"Bracket7"}
  else if (x >= 45 & x <= 49){"Bracket8"}
  else {"Bracket9"}
}

bracket_creator <- Vectorize(bracket_creator)

add_age_bracks <- function(df){
  df %>% 
    mutate_at("age", list(brackets = bracket_creator)) %>% 
    select("age", brackets, everything())
}

#Verify ages are within set
personas%>% 
  get_scores() %>% 
  norm_by_var(dimensions) %>% 
  add_age_bracks() %>% 
  assert(in_set("Bracket1","Bracket2","Bracket3", "Bracket4", "Bracket5", "Bracket6", "Bracket7","Bracket8", "Bracket9"), brackets)
  

```
# [MISSING FILL THIS IN]
# Standardizing scores based on age brackets
This will give us a better estimation of where each person stands in those dimension with respect to scores of those around their age. 
```{r Standardizing Scores Based on Age Brackets}
#std_by_bracket <- 
#[Code here for normalized function]

#[Code here that verifies normalized values created are within the bounds of 0 and 1]

```

# Simplifying Neuroticism
Currently neurtocism score is just a number and it would be helpful to quantify based on if the score is high low or average. If it is within the range of 0.5 and -0.5 then it means that the score is average neuroticism

```{r}

norm_by_brack_simple <- function(x){
  if (x > 0.3){"High"}
  else if (x < -0.3){"Low"}
  else {"Average"}
}

norm_by_brack_simple <- Vectorize(norm_by_brack_simple)

simple_neuro <- function(df){
  df %>% 
    mutate_at("neuroticism_norm_std_brack", list(neuroticism_bracks_simple = norm_by_brack_simple)) %>% 
    select(neuroticism_bracks_simple, everything())
}

#verifying all scores got assigned a value
personas%>% 
  get_scores() %>% 
  norm_by_var(c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")) %>% 
  add_age_bracks() %>% 
  mean_by_bracket(c("extraversion_norm",
                    "neuroticism_norm", 
                    "agreeableness_norm",
                    "conscientiousness_norm", 
                    "openness_norm")) %>% 
  simple_neuro()
  

``` 


# Further Work
We can go deeper into these and do the same thing with gender.


# Gender Definitions
This is going to go a little bit deeper and help us compare the scores based on age brackets as well as gender. 

Let's start by creating a new column called `ngender` which defines the genders as:
1=Male, 2=Female, 3=Other (0=missed)
```{r}
gendrify <- function(x){
  if (x == 1){"Male"}
  else if (x == 2){"Female"}
  else if (x == 3){"Other"}
  else {"Missed"}
}

gendrify <- Vectorize(gendrify)

add_gender_defs <- function(df){
  df %>% 
    mutate_at("gender", list(gender_defined = gendrify)) %>% 
    select("gender", gender_defined, everything())
}

#Checking if the any values were missed.

personas %>% 
  add_gender_defs() %>% 
  assert(in_set("Male","Female","Other","Missed"), gender_defined)

```

We can remove all the values that are"Missed" from our data as it will not add anything to the analysis
```{r}
personas %>% 
  add_gender_defs() %>% 
  filter(gender_defined != "Missed")
```

#Standardize by age bracket and then by gender
We can now find the dimensional scores by first standardizing by age and then standardizing by gender. Both can be used to see if either of those factors have any affect on how a person scores on neuroticism.
```{r}
scale2 <- function(x, na.rm = T)(x -mean(x, na.rm = na.rm))/sd(x, na.rm = na.rm)

mean_by_gender <- function(df, vars){
  df %>% 
    group_by(gender) %>% 
    mutate_at(vars, list(std_gender = scale2)) %>% 
    ungroup() %>% 
    select(age, brackets, contains('std_gender'), everything())
}

dimensions <- c("extraversion","neuroticism", "agreeableness","conscientiousness", "openness")
dimensions_std_gender <- c("extraversion_std_gender",
                           "neuroticism_std_gender", 
                           "agreeableness_std_gender",
                           "conscientiousness_std_gender", 
                           "openness_std_gender")

#Verifying the new column with scores standardized against gender do not have any NA's
personas%>% 
  get_scores() %>% 
  norm_by_var(dimensions) %>% 
  add_age_bracks() %>% 
  mean_by_bracket(dimensions) %>% 
  add_gender_defs() %>% 
  filter(gender_defined != "Missed") %>% 
  mean_by_gender(dimensions) %>% 
  verify(!is.na(dimensions_std_gender))

```

#Explanations
We can now see how each individual ranks in scores of ectraversion, agreeableness, openness and conscientiosness, based on their age bracked and their gender. We can then see how those scores have any effect on how that individual scores in neuroticism. We can now test and see if a model to predict neuroticism using these factors would be viable. 
